A Hierarchical Bayesian Framework for Exoplanetary Stellar Obliquity Distribution

### 2023-01-20 src/tex/ms.tex
[showyourwork] overleaf sync

\newcommand{\numistar}{62\xspace}
\newcommand{\numall}{161\xspace}

@@

Stellar obliquity $\psi$ describes the angle between a planet's orbital axis $\bf{n}_{\rm orb}$ and its host star's spin axis $\bf{n}_{\star}$. 
The angle is an important tracer of the formation environment and dynamical evolution of a planetary system. Stellar obliquity evolution can be roughly summarized into three stages. First, the formation and evolution of a protoplanetary disk around its host star determine the primordial stellar obliquity \citep[e.g.,][]{Bate10, Lai11, Batygin12}. Second, post-formation dynamical evolution in the planetary system, such as planet-planet scattering \citep[e.g.,][]{Rasio96, Chatterjee08, Nagasawa08, Beague12} and von Zeipel-Kozai-Lidov mechanisms \citep[e.g.,][]{Wu03, Naoz16}, excites the mutual inclinations between planetary or stellar companions, further sculpting the stellar obliquity. Lastly, tidal realignment of the host star's spin axis to the planet's orbital axis could reduce the stellar obliquity, if the stellar tidal dissipation is efficient \citep[e.g.,][]{Winn10, Albrecht12}. Furthermore, massive stars with convective cores could generate internal gravity waves and dissipate angular momentum to their radiative zones, thus modifying stellar obliquities \citep{Rogers12, Rogers13}.

It is yet unclear which one or more of these mechanisms predominantly sculpts the stellar obliquities in exoplanetary systems and what is the linkage between stellar obliquities and the formation pathways of close-in planets.
For aligned systems, because of tidal realignment, they could have different stellar obliquities in the past, making it hard to trace the obliquity-sculpting mechanism.
Misaligned systems with stellar obliquities that have not yet been washed out by tidal effects, on the other hand, still provide clues on obliquity-sculpting mechanisms.
Proposed mechanisms have different predictions on the stellar obliquity distributions. For example, in the secular chaos mechanism, stellar obliquity (or more accurately mutual inclinations between planets) could never get beyond $90\degr$ \citep{Teyssandier19}; in the stellar Kozai mechanism, stellar obliquity distribution should peak at $40\degr$ and $140\degr$ \citep{Anderson16}.
Motivated by these various predictions, we build a statistical model to characterize the stellar obliquity distribution of exoplanetary systems to distinguish between proposed stellar-obliquity-sculpting mechanisms.

@@

To find the unknown stellar obliquity distribution of exoplanetary systems, we build a hierarchical Bayesian framework that takes sky-projected stellar obliquities as observed data. In Section~\ref{sec:jacobian}, we found that the stellar inclination is not sensitive to the stellar obliquity distribution. Therefore, we focus on a flexible framework that makes the stellar inclination information optional.

\begin{figure}[ht!]
    \script{graph.py}
    \includegraphics[width=\linewidth]{figures/graph.pdf}
    \caption{Hierarchical Bayesian framework to infer the stellar obliquity distribution of exoplanetary systems. The stellar obliquity distribution is described by hyperparameters $\vb*{\beta}$ and constrained by $n_{\rm th}$ system's stellar obliquity $\psi_n$. Each $\psi_n$ is calculated by the system's sky-projected stellar obliquity $\lambda_n$ and stellar inclination $i_{\star, n}$, if available. $\gamma_{\star, n}$ contains properties of the star other than inclination, such as stellar radius and rotation period for $i_{\star, n}$ inference. Obs$_{\star,n}$ includes all the observed properties of the star with uncertainties, and $\hat{\lambda}_n$ has the measured sky-projected stellar obliquity and its uncertainty.}
    \label{fig:graph}
\end{figure}

Figure~\ref{fig:graph} shows the probabilistic graphical model for the hierarchical Bayesian framework. We aim to constrain a set of hyperparameters $\bm{\beta}$ that describe the stellar obliquity $\psi$ distribution. The parameter set $\bm{\beta}$ is constrained by $N$ individual systems in which $\psi_n$ is a random variable. $\psi_n$ is computed from the sky-projected stellar obliquity $\lambda_n$ and the stellar inclination $i_n$. 
For simplicity, we approximate $i_{\rm orb}$ to $90\degr$ and use the simplified version of Equation~(\ref{eqn:psi}) to compute $\psi_n$. $\psi$ could be computed more accurately by including $i_{\rm orb}$ as another random variable to be constrained by e.g., transit observations.
The parameter $\gamma_{\star, n}$ is another parameter set that contains all the stellar properties other than $i_\star$, e.g., stellar rotation period $P_{\rm rot}$ and radius $R_\star$, if the information is available.
Obs$_{\star,n}$ includes observed data that helps to constrain $i_{\star,n}$ and $\gamma_{\star, n}$.
If the rotational modulation method is used to constrain the stellar inclination, the $i_{\star,n}$ and $\gamma_{\star,n}$ are constrained by rotational modulation in photometric or spectroscopic time series and observed sky-projected stellar rotational line broadening $\hat{v}\sin{i_{\star,n}}$ \citep[e.g.,][]{Masuda20}, where we have $v\sin{i_\star} = 2 \pi R_\star / P_{\rm rot}$.
If the gravity-darkening method is used, the $i_{\star,n}$ is constrained by the anomaly in transit light curves. If asteroseismology is used, the $i_{\star,n}$ is constrained from the periodic variation in the photometric time series.
Lastly, the sky-projected stellar obliquity $\lambda_n$ is constrained by observed $\hat{\lambda}_n$ via the Rossiter-McLaughlin effect or gravity darkening.

For the stellar obliquity distribution, we model the $\cos{\psi}$ distribution instead of $\psi$ distribution to aim to understand if the stellar obliquity is isotropically distributed or not. We adopt a two-component mixture Beta distribution with hyperparameters $\bm{\beta} = \{\bm{w},\bm{a},\bm{b}\}$, where each hyperparameter has a dimension of 2. One component is designed for a large population of well-aligned systems, and the other component is designed for misaligned systems. The Beta distribution is a flexible distribution that is able to describe the spike of aligned systems with stellar obliquities close to $0\degr$ and also the clustered or broadly distributed misaligned systems.
The probability density function of $\cos{\psi}$ follows 
\begin{align}
    \cos{\psi} &\sim 2\times\Bigl( w_0 {\rm Beta}(a_0, b_0) + w_1 {\rm Beta}(a_1, b_1)\Bigr)-1.
\end{align}
Since the Beta distribution is defined on the interval $[0, 1]$, we extend its support to $[-1,1]$ using a linear transformation (i.e., $2\times {\rm Beta} - 1$). For the ${\rm Beta}(a_0,b_0)$ component, we have it to describe the population of aligned systems. For the $a_0/b_0 \gg 1$, the distribution will have a spike at $\cos{\psi} = 1$. For the ${\rm Beta}(a_1,b_1)$ component, we have it to describe the misaligned systems. We design the order of the two components to avoid label switching in the mixture model.
With the design, the hyperparameter and parameter priors are the following:
\begin{align}
    w_{0,1} &\sim {\rm Dirichlet}(1, 1) \nonumber\\
    a_0 &\sim \mathcal{U}(0, 50) \nonumber\\
    b_0 &\sim \mathcal{U}(0, 1) \nonumber\\
    a_1 &\sim \mathcal{U}(0, 10) \nonumber\\
    b_1 &\sim \mathcal{U}(0, 10) \nonumber\\
    \cos{i}_{\star,n} &\sim \mathcal{U}(0, 1) \nonumber\\
    \lambda_{\star,n} &\sim \mathcal{U}(0, \pi).
\end{align}
The likelihood functions follow:
\begin{align}
    \mathcal{L}(\lambda) &\sim \prod_{i=1}^N\mathcal{N}(\hat{\lambda}_n, \sigma_{\hat{\lambda}_n}) \nonumber\\
    \mathcal{L}(i_\star) &\sim \prod_{i=1}^N\mathcal{N}({\rm Obs}_{\star,n}, \sigma_{{\rm Obs}_{\star,n}}).
\end{align}
If $\gamma_{\star, n}$ is available, we construct Normal distributions with means and standard deviations from their observed data.
Here we assume uniform hyperpriors for the Beta distribution. For population inference on a small number of systems, the choice of hyperpriors could impact the inferred distribution \citep{Nagpal22}. When applying the framework to a small sample size, it is important to use different hyperpriors to test the robustness of the model.

The probabilistic model is built with the $\mathtt{PyMC}$ \citep[$\mathtt{v4.1.7}$;][]{pymc} package and the posteriors are sampled using the No-U-Turn Sampler \citep[NUTS;][]{Hoffman11}, a gradient-based sampling algorithm of the Markov chain Monte Carlo (MCMC). All figures and simulations in this paper are fully reproducible and built with the $\mathtt{showyourwork}$ package. The source code can be found on \href{https://github.com/jiayindong/polar}{GitHub\,\faGithub}.

@@

We first apply the hierarchical Bayesian framework to simulated data of which the ground-truth stellar obliquity distribution is known. We adopt the four $\cos{\psi}$ distributions discussed in Section~\ref{sec:jacobian}: a uniform distribution $\mathcal{U}(-1,1)$ and three Normal distributions $\mathcal{N}(0,0.2)$, $\mathcal{N}(-0.4,0.2)$, and $\mathcal{N}(0.4,0.2)$.
For each $\cos{\psi}$ distribution, we randomly generate 200 samples of stellar inclination $i_\star$ and sky-projected stellar obliquity $\lambda$, assuming a uniform $\theta$. The sampled $i_\star$ and $\lambda$ here are \emph{true} values. We then add some Gaussian noises to the \emph{true} $\lambda$ and $i_\star$ to simulate the \emph{observed} data. We choose uncertainties of $\sigma_{\lambda} = 8\degr$ and $\sigma_{i_\star} = 10\degr$, which are typical observational uncertainties summarized in \cite{Albrecht22}. Using the observed $\lambda$ and their uncertainties, we infer the $\cos{\psi}$ distribution of the sample with or without $i_\star$ likelihoods.

In Figure~\ref{fig:simulation}, we present the inferred stellar obliquity distributions. Since the simulated stellar obliquity distributions only have a single component, we model the data with a single Beta distribution. Each row corresponds to an injected stellar obliquity distribution. The orange curve and contours are the median and uncertainties of the inferred $\cos{\psi}$ distribution with stellar inclination information, and the blue curve and contours are the ones without stellar inclination information. We simply assume an isotropic distribution of $i_\star$ for these models.
As shown in Figure~\ref{fig:simulation}, in all four distributions, the inferred $\cos{\psi}$ distributions with or without $i_\star$ correctly recover the injected distributions shown as grey dashed lines. Since some of the injected distributions are Normal distributions, it is expected that the inferred distributions, which as Beta distributions, could not exactly describe the injected distributions.
Comparing the blue curves with the orange curves, the overestimation of $\cos{\psi}$ near $\pm1$ under the assumption of an isotropic $i_\star$ has little impact on the inference. The distribution is mainly constrained by the peak and width of the stellar obliquity distribution, which are both correctly predicted by the isotropic $i_\star$ model.

From the simulated data, we confirm that while stellar inclinations could help to constrain the stellar obliquity of individual systems, they are not necessarily required for the population stellar obliquity inference. The constraint on stellar obliquity distribution is predominantly based on sky-projected stellar obliquities. The sky-projected stellar obliquity could correctly infer the peak and width of the underlying stellar obliquity distribution.

@@

We then apply the hierarchical Bayesian framework to all exoplanetary systems with sky-projected stellar obliquity measurements. We use the sample of \numall systems with sky-projected stellar obliquity $\lambda$ measurements summarized in \cite{Albrecht22} Table A1, which are predominantly Hot Jupiter systems. The inferred $\cos{\psi}$ is shown in Figure~\ref{fig:psi_dist}. The $\cos{\psi}$ distribution peaks at 1 but is nearly flat between $-0.75$ and $0.75$ with no significant clustering. The distribution corresponds to an isotropic distribution of $\psi$ from $40\degr$ to $140\degr$. 
The posteriors of the hyperparameters are \textcolor{red}{to be added}.

@@

The inferred stellar obliquity distribution of the full sample of systems is at odds with the previous analysis of the subsample of systems that also have $i_\star$ measurements. The subsample identified a preponderance of perpendicular planets that disfavored an isotropic stellar obliquity distribution \citep{Albrecht21}. We summarize two possible reasons to explain the discrepancies, which are open to further investigation.
First, the sample of systems with $i_\star$ measurements is small, and only about 20 systems are misaligned systems used to constrain the misaligned systems' stellar obliquity distribution. Second, the sample could be biased because of the $i_\star$ measurements.

Our analysis suggests misaligned systems have nearly isotropic stellar obliquity distribution. When comparing the inferred distribution with the prediction of stellar obliquity distributions in different Hot Jupiter origin scenarios, it is in best agreement with the outcome of the scattering of multiple giant planets after a convergent disk migration \citep{Beague12}.

\section{Transformation between the $\psi$, $\lambda$, and \lowercase{$i_\star$} distributions}\label{sec:jacobian}

In this section, we aim to understand why the stellar obliquity distribution is predominantly determined by the sky-projected stellar obliquity distribution and why the stellar inclination distribution is less important to the inference.

@@

Stellar obliquity $\psi$ is the angle between a planet's orbital axis $\bf{n}_{\rm orb}$ and its host star's spin axis $\bf{n}_{\star}$. In Figure~\ref{fig:coord}, we build two coordinate systems to describe the stellar spin axis $\bf{n}_{\star}$ for a given orbital axis $\bf{n}_{\rm orb}$. In both coordinates, we set the observer's line of sight to be one of the two horizontal axes, and the orbital axis of the planet to be the vertical axis. We approximate the orbital inclination of the transiting planet to $90\degr$ here to simplify the problem for a population study.

@@

Due to the lack of dependency on the $i_\star$ distribution, the $\psi$ distribution could be inferred purely from the $\lambda$ distribution without losing much information. Although we could find a relation between $\psi$, $\lambda$, and $\theta$, the $\psi$ distribution cannot be inferred from the $\lambda$ and $\theta$ distributions since $\lambda$ and $\theta$ are not independent variables. We still need to infer the $\psi$ distribution from the $\lambda$ and $i_\star$ distributions.
In the \nth{4} column in Figure~\ref{fig:transform}, we find the $\psi$ distribution from the $\lambda$ distribution, simply assuming an isotropic $i_\star$ distribution. As shown in grey histograms, the method could correctly recover the peaks and widths of the underlying $\psi$ distributions (presented as blue dashed lines). The method over-predicts $\cos{\psi}$ at $\pm 1$ because of the over-prediction of $i_\star$ near $90\degr$. However, this minor deviation from the true distribution has little impact on the population inference because of hyperparameters' limited degree of freedom.

@@

In this work, we demonstrate the stellar obliquity distribution inference is predominantly based on the sky-projected stellar obliquity. While stellar inclination improves the stellar obliquity constraint on individual systems, it does not improve the population-level stellar obliquity constraint. 
We introduce a flexible, hierarchical Bayesian framework to infer the stellar obliquity distribution of a population. The model is an open-source code available to the public to apply to a certain sample of targets.
When applying our framework to all exoplanetary systems with sky-projected stellar obliquity measurements, we found no significant clustering near $\cos{\psi} = 0$, at odds with the previous study of systems that have both sky-projected stellar obliquity and stellar obliquity measurements. Potential biases that lead to this discrepancy are open to discussion. 
Their stellar obliquity is nearly isotropically distributed for misaligned exoplanetary systems, which may suggest a multi-planet, convergent-disk-migration history followed by a scattering phase in these planets' dynamical history. Future studies on a more carefully selected Hot Jupiter sample will put better constraints on the origin of Hot Jupiters from their stellar obliquity distribution.

>>>>

@@

Stellar obliquity $\psi$ describes the angle between a planet's orbital axis $\bf{n}_{\rm orb}$ and its host star's spin axis $\bf{n}_{\star}$. 
The angle is an important tracer of a planetary system's formation environment and dynamical evolution. Stellar obliquity evolution can be roughly summarized into three stages. First, the formation and evolution of a protoplanetary disk around its host star determine the primordial stellar obliquity \citep[e.g.,][]{Bate10, Lai11, Batygin12}. Second, post-formation dynamical evolution in the planetary system, such as planet-planet scattering \citep[e.g.,][]{Rasio96, Chatterjee08, Nagasawa08, Beague12} and von Zeipel-Kozai-Lidov mechanisms \citep[e.g.,][]{Wu03, Naoz16}, excites the mutual inclinations between planetary or stellar companions, further sculpting the stellar obliquity. Lastly, tidal realignment of the host star's spin axis to the planet's orbital axis could reduce the stellar obliquity, if the stellar tidal dissipation is efficient \citep[e.g.,][]{Winn10, Albrecht12}. Furthermore, massive stars with convective cores could generate internal gravity waves and dissipate angular momentum to their radiative zones, thus modifying stellar obliquities \citep{Rogers12, Rogers13}.

It is yet unclear which one or more of these mechanisms predominantly sculpts the stellar obliquities in exoplanetary systems and the linkage between stellar obliquities and the formation pathways of close-in planets.
Because of tidal realignment, aligned systems could have had different stellar obliquities in the past, making it hard to trace the obliquity-sculpting mechanism.
On the other hand, misaligned systems with stellar obliquities that have not yet been washed out by tidal effects still provide clues on obliquity-sculpting mechanisms.
Proposed mechanisms have different predictions on the stellar obliquity distributions. For example, in the secular chaos mechanism, stellar obliquity (or more accurately mutual inclinations between planets) could never get beyond $90\degr$ \citep{Teyssandier19}; in the stellar Kozai mechanism, stellar obliquity distribution should peak at $40\degr$ and $140\degr$ \citep{Anderson16}.
Motivated by these predictions, we build a statistical model to characterize the stellar obliquity distribution of exoplanetary systems to distinguish between proposed stellar-obliquity-sculpting mechanisms.

@@

To find the unknown stellar obliquity distribution of exoplanetary systems, we build a hierarchical Bayesian framework that takes the sky-projected stellar obliquity $\lambda$ and orbital inclination $i_{\rm orb}$ as observed data. Stellar inclination $i_\star$ is an optional and flexible input. It could be directly taken from previous observations, or if the rotational modulation technique is used, the stellar inclination will be inferred from the stellar rotation period $P_{\rm rot}$, stellar radius $R_\star$, and sky-projected rotational broadening velocity $v\sin{i}_\star$. If the stellar inclination information is not provided, we simply assume an isotropic stellar inclination. As we will show in later sections, assuming an isotropic stellar inclination will not compromise the stellar obliquity distribution inference.

\begin{figure}[ht!]
    \script{graph.py}
    \includegraphics[width=\linewidth]{figures/graph.pdf}
    \caption{Hierarchical Bayesian framework to infer the stellar obliquity distribution of exoplanetary systems. The stellar obliquity distribution is described by hyperparameters $\vb*{\beta}$ and constrained by $n_{\rm th}$ system's stellar obliquity $\psi_n$. Each $\psi_n$ is calculated by the system's sky-projected stellar obliquity $\lambda_n$, orbital inclination $i_{{\rm orb}, n}$, and stellar inclination $i_{\star, n}$, if available. $\gamma_{\star, n}$ contains properties of the star other than inclination, such as stellar radius and rotation period for $i_{\star, n}$ inference. Obs$_{\star,n}$ includes all the observed properties of the star with uncertainties, $\hat{\lambda}_n$ has the measured sky-projected stellar obliquity and its uncertainty, and $\hat{i}_{{\rm orb}, n}$ has measured orbital inclination and its uncertainty.}
    \label{fig:graph}
\end{figure}

Figure~\ref{fig:graph} shows the probabilistic graphical model for the hierarchical Bayesian framework. We aim to constrain a set of hyperparameters $\bm{\beta}$ that describe the stellar obliquity $\psi$ distribution. The parameter set $\bm{\beta}$ is constrained by $N$ individual systems in which $\psi_n$ is a random variable. $\psi_n$ is computed from the sky-projected stellar obliquity $\lambda_n$, the orbital inclination $i_{{\rm orb},n}$, and the stellar inclination $i_{\star, n}$. 
The parameter $\gamma_{\star, n}$ is a parameter set that contains all the stellar properties other than $i_\star$, e.g., stellar rotation period $P_{\rm rot}$ and radius $R_\star$, if the information is available.
Obs$_{\star,n}$ includes observed data that helps to constrain $i_{\star,n}$ and $\gamma_{\star, n}$.
If the rotational modulation method is used to constrain the stellar inclination, the $i_{\star,n}$ and $\gamma_{\star,n}$ are constrained by rotational modulation in photometric or spectroscopic time series and observed sky-projected stellar rotational line broadening $\hat{v}\sin{i_{\star,n}}$ \citep[e.g.,][]{Masuda20}, where we have $v\sin{i_\star} = 2 \pi R_\star / P_{\rm rot}$.
If the gravity-darkening method is used, the $i_{\star,n}$ is constrained by the anomaly in transit light curves. If asteroseismology is used, the $i_{\star,n}$ is constrained from the periodic variation in the photometric time series.
The sky-projected stellar obliquity $\lambda_n$ can be constrained by observed $\hat{\lambda}_n$ via the Rossiter-McLaughlin effect or gravity darkening.
Lastly, the orbital inclination $i_{{\rm orb},n}$ can be constrained from the transit light curves.

For the stellar obliquity distribution, we model the $\cos{\psi}$ distribution instead of $\psi$ distribution to understand if the stellar obliquity is isotropically distributed. If stellar obliquities of misaligned systems follow an isotropic distribution, the $\cos{\psi}$ distribution will be uniformly distributed between -1 to 1. We adopt a two-component mixture Beta distribution with hyperparameters $\bm{\beta} = \{\bm{w},\bm{a},\bm{b}\}$, where each hyperparameter has a dimension of 2. One component is designed for a large population of well-aligned systems, and the other is designed for misaligned systems. The Beta distribution is a flexible distribution that can describe the spike of aligned systems with stellar obliquities close to $0\degr$ or the clustered or broadly distributed misaligned systems.
The probability density function of $\cos{\psi}$ follows 
\begin{align}
    \cos{\psi} &\sim 2\times\Bigl( w_0 {\rm Beta}(a_0, b_0) + w_1 {\rm Beta}(a_1, b_1)\Bigr)-1.
\end{align}
Since the Beta distribution is defined on the interval $[0, 1]$, we extend its support to $[-1,1]$ using a linear transformation (i.e., $2\times {\rm Beta} - 1$). For the ${\rm Beta}(a_0,b_0)$ component, we have it to describe the population of aligned systems. For the $a_0/b_0 \gg 1$, the distribution will have a spike at $\cos{\psi} = 1$. For the ${\rm Beta}(a_1,b_1)$ component, we have it to describe the misaligned systems. We design the order of the two components to avoid label switching in the mixture model.
With the design, the hyperparameter and parameter priors are the following:
\begin{align}
    w_{0,1} &\sim {\rm Dirichlet}(1, 1) \nonumber\\
    a_0 &\sim \mathcal{U}(0, 50) \nonumber\\
    b_0 &\sim \mathcal{U}(0, 1) \nonumber\\
    a_1 &\sim \mathcal{U}(0, 10) \nonumber\\
    b_1 &\sim \mathcal{U}(0, 10) \nonumber\\
    \cos{i}_{\star,n} &\sim \mathcal{U}(0, 1) \nonumber\\
    \cos{i}_{{\rm orb},n} &\sim \mathcal{U}(-1, 1) \nonumber\\
    \lambda_{\star,n} &\sim \mathcal{U}(0, \pi).
\end{align}
The likelihood functions follow:
\begin{align}
    \mathcal{L}(\lambda) &\sim \prod_{i=1}^N\mathcal{N}(\hat{\lambda}_n, \sigma_{\hat{\lambda}_n}) \nonumber\\
    \mathcal{L}(i_\star) &\sim \prod_{i=1}^N\mathcal{N}({\rm Obs}_{\star,n}, \sigma_{{\rm Obs}_{\star,n}}).
\end{align}
If $\gamma_{\star, n}$ is available, we construct Normal distributions with means and standard deviations from their observed data.
Here we assume uniform hyperpriors for the Beta distribution. For population inference on a small number of systems (i.e., $N<50$), the choice of hyperpriors could impact the inferred distribution \citep{Nagpal22}. When applying the framework to small sample size, it is important to test the robustness of inferred distributions on different hyperpriors.

The probabilistic model is built with the $\mathtt{PyMC}$ \citep[$\mathtt{v4.1.7}$;][]{pymc} package and the posteriors are sampled using the No-U-Turn Sampler \citep[NUTS;][]{Hoffman11}, a gradient-based sampling algorithm of the Markov chain Monte Carlo (MCMC). All figures and simulations in this paper are fully reproducible and built with the $\mathtt{showyourwork}$ package. The open source code can be found on \href{https://github.com/jiayindong/polar}{GitHub\,\faGithub}.

@@

We first apply the hierarchical Bayesian framework to simulated data of which the ground-truth stellar obliquity distribution is known. We test the four $\cos{\psi}$ distributions: a uniform distribution $\mathcal{U}(-1,1)$ and three Normal distributions $\mathcal{N}(0,0.2)$, $\mathcal{N}(-0.4,0.2)$, and $\mathcal{N}(0.4,0.2)$.
For each $\cos{\psi}$ distribution, we randomly generate 200 samples of stellar inclination $i_\star$ and sky-projected stellar obliquity $\lambda$. We assume the stellar spin axis is uniformly distributed around the planetary orbital axis in the azimuthal direction and the orbital inclination is $90\degr$. 
The sampled $i_\star$ and $\lambda$ here are \emph{true} values. We then add some Gaussian noises to the \emph{true} $\lambda$ and $i_\star$ to simulate the \emph{observed} data. We choose uncertainties of $\sigma_{\lambda} = 8\degr$ and $\sigma_{i_\star} = 10\degr$, which are typical observational uncertainties summarized in \cite{Albrecht22}. Using the observed $\lambda$ and their uncertainties, we infer the $\cos{\psi}$ distribution of the sample with or without $i_\star$ likelihoods.

In Figure~\ref{fig:simulation}, we present the inferred stellar obliquity distributions. Since the simulated stellar obliquity distributions only have a single component, we model the data with a single Beta distribution. Each row corresponds to an injected stellar obliquity distribution. The orange curve and contours are the median and uncertainties of the inferred $\cos{\psi}$ distribution with stellar inclination information, and the blue curve and contours are the ones without stellar inclination information. We simply assume an isotropic distribution of $i_\star$ for these models.
Surprisingly, in all four distributions, the inferred $\cos{\psi}$ distributions with or without $i_\star$ correctly recover the injected distributions shown as grey dashed lines as shown in Figure~\ref{fig:simulation}. Since some of the injected distributions are Normal distributions, it is expected that the inferred distributions, which as Beta distributions, could not exactly describe the injected distributions.

We also examine the importance of orbital inclination to the stellar obliquity inference. Since all the systems we study are transiting exoplanetary systems, we consider an isotropic orbital inclination distribution from $80\degr$ to $90\degr$. This broad inclination range corresponds to an impact parameter range from 0 to 1 for a circular orbit planet with an $a/R_\star$ of 6. We find at this level of orbital inclination deviations, the difference on the inferred stellar obliquity distributions is minimal. For transiting planet systems, approximating orbital inclinations to $90\degr$ will not compromise the stellar obliquity distribution inference.

From the simulated data, we confirm that while stellar inclinations help to constrain the stellar obliquity of individual systems, they are not necessarily required for the population stellar obliquity inference. The constraint on stellar obliquity distribution is predominantly based on sky-projected stellar obliquities, and the sky-projected stellar obliquity could correctly infer the peak and width of the underlying stellar obliquity distribution.

@@

We then apply the hierarchical Bayesian framework to all exoplanetary systems with sky-projected stellar obliquity $\lambda$ measurements. We use the sample of 161 systems summarized in \cite{Albrecht22} Table A1, which are predominantly Hot Jupiter systems. The inferred $\cos{\psi}$ distribution is shown in Figure~\ref{fig:psi_dist}. The $\cos{\psi}$ distribution peaks at 1 but is nearly flat between $-0.75$ and $0.75$ with no significant clustering. The distribution corresponds to an isotropic distribution of $\psi$ from $40\degr$ to $140\degr$. 
The aligned-system population dominates the distribution with a fraction of $w_0 = 0.67 \pm 0.09$. The hyperposteriors of the Beta distribution are $a_0 = 31.0\pm12.8$ and $b_0 = 0.40\pm0.12$.
The misaligned-system population has a fraction of $w_1 = 0.33 \pm 0.09$ and hyperposteriors of $a_1 = 1.56\pm0.93$ and $b_1 = 1.64\pm1.58$.

@@

The inferred stellar obliquity distribution of the full sample of systems is at odds with the previous analysis of the subsample of systems that also have $i_\star$ measurements. The subsample identified a preponderance of perpendicular planets that disfavored an isotropic stellar obliquity distribution \citep{Albrecht21}. We speculate two possible explanations for this observation discrepancy, which are open to further investigation.
First, the sample of systems with $i_\star$ measurements is small, and only about 20 systems are misaligned systems used to constrain the misaligned systems' stellar obliquity distribution. Second, the sample could have a selection bias introduced by the $i_\star$ measurement requirement.

Our analysis suggests misaligned systems have nearly isotropic stellar obliquity distribution. When comparing the inferred distribution with the prediction of stellar obliquity distributions in different Hot Jupiter origin scenarios, it is in best agreement with the outcome of the scattering of multiple giant planets after a convergent disk migration \citep[e.g.,][]{Beague12}, which could produce planets with orbital inclination from $0\degr$ to $180\degr$ with a peak at $0\degr$. This interesting result can be further tested on more carefully selected Hot Jupiter samples and put constraints on Hot Jupiter origins.

\section{Relations between the $\psi$, $\lambda$, and \lowercase{$i_\star$} distributions}\label{sec:jacobian}

In this section, we aim to use mathematical expressions to understand why the stellar obliquity distribution is predominantly determined by the sky-projected stellar obliquity distribution and why the stellar inclination distribution is less important to the inference.

@@

In Figure~\ref{fig:coord}, we build two coordinate systems to describe the stellar spin axis $\bf{n}_{\star}$ for a given orbital axis $\bf{n}_{\rm orb}$. In both coordinates, we set the observer's line of sight to be one of the two horizontal axes, and the orbital axis of the planet to be the vertical axis. We approximate the orbital inclination of the transiting planet to $90\degr$ here to simplify the problem for a population study. We found in the previous section that the assumption would not compromise the stellar obliquity distribution inference for transiting planets.

@@

Due to the lack of dependency on the $i_\star$ distribution, the $\psi$ distribution could be inferred purely from the $\lambda$ distribution without losing much information. Although we could find a relation between $\psi$, $\lambda$, and $\theta$, the $\psi$ distribution cannot be inferred from the $\lambda$ and $\theta$ distributions since $\lambda$ and $\theta$ are not independent variables. We still need to infer the $\psi$ distribution from the $\lambda$ and $i_\star$ distributions.
In the \nth{4} column in Figure~\ref{fig:transform}, we find the $\psi$ distribution from the $\lambda$ distribution, simply assuming an isotropic $i_\star$ distribution. The method could correctly recover the peaks and widths of the underlying $\psi$ distributions (presented as blue dashed lines) as shown in grey histograms. The method over-predicts $\cos{\psi}$ at $\pm 1$ because of the over-prediction of $i_\star$ near $90\degr$. However, this minor deviation from the true distribution has little impact on the population inference because of hyperparameters' limited degree of freedom. Comparing the blue curves with the orange curves in Figure~\ref{fig:simulation}, the overestimation of $\cos{\psi}$ near $\pm1$ under the assumption of an isotropic $i_\star$ has little impact on the inference since the distribution is mainly constrained by the peak and width of the stellar obliquity distribution, which are both correctly predicted by the isotropic $i_\star$ model.

@@

In this work, we demonstrate the stellar obliquity distribution inference is predominantly based on the sky-projected stellar obliquity. While stellar inclination improves the stellar obliquity constraint on individual systems, it does not improve the population-level stellar obliquity constraint. 
We introduce a flexible, hierarchical Bayesian framework to infer the stellar obliquity distribution of a population. Information on stellar inclination is not required for the stellar obliquity distribution inference; if the stellar inclination is not provided, we simply assume an isotropic stellar inclination distribution. When combining samples with and without stellar inclination measurements, it is important to check if the $i_\star$ sample is representative of the full sample. If not, since the $i_\star$ sample will constrain stellar obliquities more tightly, the overall distribution could be weighted towards these systems, bias our interpretation.
The hierarchical Bayesian model we introduce in this paper is an open-source code available to the public to apply to a certain sample of targets and customize for different stellar obliquity distributions and priors.

When applying our framework to all exoplanetary systems with sky-projected stellar obliquity measurements, we found no significant clustering near $\cos{\psi} = 0$, at odds with the previous study of systems that have both sky-projected stellar obliquity and stellar obliquity measurements. Potential biases that lead to this discrepancy are open to discussion. 
Their stellar obliquity is nearly isotropically distributed for misaligned exoplanetary systems in support of a multi-planet, convergent-disk-migration history followed by a scattering phase in these planets' dynamical history. Future studies on a more carefully selected Hot Jupiter sample will put better constraints on the origin of Hot Jupiters from their stellar obliquity distribution.
